---
params:
  subtitle: ""
  title: ""
  authors: ""
title: '`r params$title`'
subtitle: '`r params$subtitle`'
author: '`r params$authors`'
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    self_contained: false
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: TRUE
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
  bookdown::word_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
    fig_width: 5
always_allow_html: yes
bibliography: '`r paste0(here::here(), "/bibliography.bib")`'
---

```{r knitrSetup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # by default turn off code echo
```


```{r codeSetup, include=FALSE}
# Set start time ----
startTime <- proc.time()
# Local parameters ----
b2Kb <- 1024 #http://whatsabyte.com/P1/byteconverter.htm
b2Mb <- 1048576
plotLoc <- here::here("docs","plots") # where to put the plots

library(gridCarbon)
# Packages used in the report ----
rmdLibs <- c("ggplot2", # plots
          "kableExtra", # fancy tables
          #"openair" # air qual plots re-purposed, fails as needs latticeExtra which needs R (≥ 3.6.0) https://cran.r-project.org/web/packages/latticeExtra/index.html
          "lubridate" # dates
          )
gridCarbon::loadLibraries(rmdLibs)

# Local parameters ----
localParams$plotPath <- here::here("docs", "plots/")
```

# Get national demand data

```{r gridGen}

allGridDT <- drake::readd(gridGenData)

MWhDT <- allGridDT[, .(MWh = sum(kWh/1000)), keyby = .(rDateTimeNZT, rTime, season, year)]
```

# Heat pump impact simple simulation

Create an animation of a simulation of flattening peak. 

We use GREEN Grid heat pump data to calculate the mean kW per half hour.

```{r loadHeatPump}

hpDT <- drake::readd(heatPumpData)

h <- head(hpDT[,.(r_dateTime, r_dateTimeLubridated, rDateTimeNZT, halfHour)])
h

#uniqueN(hpDT$halfHour)
hpDT[, .(minTime = min(hms),
         maxTime = max(hms)),
     keyby = .(halfHour)]


baselineDT <- hpDT[, .(meankW = mean(powerW/1000)), keyby = .(year, season, halfHour)]


# check plot
ggplot2::ggplot(baselineDT[year == 2015], aes(x = halfHour, y = meankW, colour = season)) +
  geom_line() +
  facet_grid(year ~ .) +
  labs(x = "Time",
       y = "Mean kW",
       caption = "Data: NZ GREEN Grid households")
```

which suggests ~0.6 kW power in the morning peak and 0.5 in the evening. Perhaps people light woodburners in the evening?

BRANZ HCS 2015:
 * 40% owner-occupiers have HPs
 * 25% rentals

NZ Census:
 * 64% X m owner-occupied -> 1,228,500
 * 32% Y m rentals -> 609,700
 * 4% (?) rent-free -> 65,200 (we treat these as owned)

https://www.stats.govt.nz/information-releases/dwelling-and-household-estimates-september-2019-quarter


First we inflate it by adding 1 kW heat pump per house - which we assume needs 0.5 kWh per half hour. For 2 million households. That +1 GWh :-)

```{r kWhWinterSimulation}
nOwner <- 1228500 + 65200 # owned + rent free
nRenter <- 609700 # rented
hpsOwner <- 0.4 * nOwner
hpsRenter <- 0.25 * nRenter



baselineDT[, hpCurrentMeanMW := ((meankW * hpsOwner) + (meankW * hpsRenter))/1000]
baselineDT[, hpCurrentMeanMWh := hpCurrentMeanMW/2]

# model to 100%
newHPsOwner <- nOwner * (1-0.4) # 40% already have them
newHPsRenter <- nRenter * (1-0.25) # 40% already have them

baselineDT[, hpFutureMeanMW := ((meankW * nOwner) + (meankW * nRenter))/1000]
baselineDT[, hpFutureMeanMWh := hpFutureMeanMW/2]

p <- ggplot2::ggplot(baselineDT[year == 2015], aes(x = halfHour, y = hpCurrentMeanMWh, colour = season)) +
  geom_step() +
  theme(legend.position="bottom") +
  scale_color_discrete(name="Season") +
  labs(x = "Time",
       y = "Mean MWh (heat pumps)",
       caption = "Year: 2019\nData source: NZ GREEN Grid")

p
```

```{r buildHPModel}
MWhDT[, halfHour := hms::trunc_hms(rTime, 30*60)]
gridDT <- MWhDT[, .(meanMWh = mean(MWh)), keyby = .(halfHour, season, year)]

ggplot2::ggplot(gridDT[year == 2017], aes(x = halfHour, y = meanMWh, colour = season)) +
  geom_line() +
  facet_grid(year ~ .) +
  labs(x = "Time",
       y = "MWh (total)",
       caption = "Data: NZ Electricity Authority")

# just take winter
winter2017DT <- gridDT[year == 2017 & season == "Winter"]
winter2017DT[, model := "2017 generation"]
setkey(winter2017DT, halfHour, season, year)

winterHpCurrentDT <- baselineDT[year == 2015 & season == "Winter", .(year, meanMWh = hpCurrentMeanMWh, 
                                                                     halfHour, season)]
winterHpCurrentDT[, year := 2017] # fix
winterHpCurrentDT[, model := "Modelled Heat pump (2017)"]

winterHpFutureDT <- baselineDT[year == 2015 & season == "Winter", .(year, hpFutureMeanMWh, 
                                                                     halfHour, season)]
winterHpFutureDT[, year := 2017] # fix
setkey(winterHpFutureDT, halfHour, season, year)
winterHpFutureDT <- winterHpFutureDT[winter2017DT]
winterHpFutureDT[, meanMWh := hpFutureMeanMWh + meanMWh]
winterHpFutureDT[, model := "100% heat pump"]

modelDT <- rbind(winter2017DT[, .(year, halfHour, season, meanMWh, model)],
                 winterHpCurrentDT[, .(year, halfHour, season, meanMWh, model)],
                 winterHpFutureDT[, .(year, halfHour, season, meanMWh, model)])
                 
ggplot2::ggplot(modelDT[year == 2017], aes(x = halfHour, y = meanMWh/1000, colour = model)) +
  geom_line() +
  facet_grid(year ~ season) +
  scale_color_discrete(name="Model") +
  labs(x = "Time",
       y = "GWh (total)",
       caption = "Data: NZ Electricity Authority, NZ GREEN Grid households, \nNZ Census 2018, BRANZ HCS 2015")

```

# Flatten the curve plots

Another simulation

```{r flattenTheCurve}

leveled20pcDT <- gridDT[year == 2017 & season == "Winter"]
overallMeanMWh <- mean(leveled20pcDT$meanMWh)
leveled20pcDT[, diff := meanMWh - overallMeanMWh]
leveled20pcDT[, levelled := (0.2 * -1 * diff) + meanMWh]
leveled20pcDT[, meanMWh := levelled]
leveled20pcDT[, model := "20% flatten"]

leveled40pcDT <- gridDT[year == 2017 & season == "Winter"]
leveled40pcDT[, diff := meanMWh - overallMeanMWh]
leveled40pcDT[, meanMWh := (0.4 * -1 * diff) + meanMWh]
leveled40pcDT[, model := "40% flatten"]

leveled60pcDT <- gridDT[year == 2017 & season == "Winter"]
leveled60pcDT[, diff := meanMWh - overallMeanMWh]
leveled60pcDT[, meanMWh := (0.6 * -1 * diff) + meanMWh]
leveled60pcDT[, model := "60% flatten"]

leveled80pcDT <- gridDT[year == 2017 & season == "Winter"]
leveled80pcDT[, diff := meanMWh - overallMeanMWh]
leveled80pcDT[, meanMWh := (0.8 * -1 * diff) + meanMWh]
leveled80pcDT[, model := "80% flatten"]

# p <- ggplot(leveledDT, aes(x = halfHour)) + 
#   geom_line(aes(y = levelled), colour = "red") +
#   geom_line(aes(y = diff), colour = "blue") +
#   geom_line(aes(y = meanMWh))

modelDT <- rbind(winter2017DT[, .(year, halfHour, season, meanMWh, model)],
                 winterHpCurrentDT[, .(year, halfHour, season, meanMWh, model)],
                 winterHpFutureDT[, .(year, halfHour, season, meanMWh, model)], 
                 leveled20pcDT[, .(year, halfHour, season, meanMWh, model)], 
                 leveled40pcDT[, .(year, halfHour, season, meanMWh, model)], 
                 leveled60pcDT[, .(year, halfHour, season, meanMWh, model)], 
                 leveled80pcDT[, .(year, halfHour, season, meanMWh, model)])

modelDT[, model := factor(model, levels = c("2017 generation",
                                              "100% heat pump",
                                              "20% flatten", 
                                              "40% flatten", 
                                              "60% flatten", 
                                              "80% flatten")
                           )
        ]
```

```{r flattenTheCurvePlots}
yMax <- max(modelDT$meanMWh/1000) # GWh
myCap <- "Data: NZ Electricity Authority, NZ GREEN Grid households, \nNZ Census 2018, BRANZ HCS 2015"

drawPlot <- function(dt){
  p <- ggplot2::ggplot(dt, 
                  aes(x = halfHour, y = meanMWh/1000,  colour = model)) +
  geom_line() +
  scale_color_discrete(name="Ambition") +
  # theme(legend.position="bottom")
  ylim(1.5,yMax) +
  labs(x = "Time",
       y = "Demand")
  p <- p + theme_dark() +
    theme(plot.background = element_rect(fill = "black"),
          legend.background = element_rect(fill = "darkGrey"))
  return(p)
}

plotDT <- modelDT[year == 2017 & !(model %like% "Modelled") & 
                               !is.na(model)]
dt <- plotDT[model %like% "gen"]
p1 <- drawPlot(dt)
p1
ggplot2::ggsave(filename = "flattenTheCurve_1.png", 
                     path = localParams$plotPath)

p2 <- drawPlot(plotDT[model %like% "gen" | 
                               model %like% "100%"])
p2
ggplot2::ggsave(filename = "flattenTheCurve_2.png", 
                     path = localParams$plotPath)

p3 <- drawPlot(plotDT[model %like% "gen" | 
                               model %like% "100%" | 
                               model %like% "20%"])
p3
ggplot2::ggsave(filename = "flattenTheCurve_3.png", 
                     path = localParams$plotPath)

p4 <- drawPlot(plotDT[model %like% "gen" | 
                               model %like% "100%" | 
                               model %like% "20%" | 
                               model %like% "40%"])
p4
ggplot2::ggsave(filename = "flattenTheCurve_4.png", 
                     path = localParams$plotPath)

p5 <- drawPlot(plotDT[model %like% "gen" | 
                               model %like% "100%" | 
                               model %like% "20%" | 
                               model %like% "40%" | 
                               model %like% "60%" ])
p5
ggplot2::ggsave(filename = "flattenTheCurve_5.png", 
                     path = localParams$plotPath)

p6 <- drawPlot(plotDT[model %like% "gen" | 
                               model %like% "100%" | 
                               model %like% "20%" | 
                               model %like% "40%" | 
                               model %like% "60%" | 
                               model %like% "80%"])
p6
ggplot2::ggsave(filename = "flattenTheCurve_6.png", 
                     path = localParams$plotPath)
```



# Runtime

```{r check runtime, include=FALSE}
t <- proc.time() - startTime
elapsed <- t[[3]]
```

Analysis completed in `r round(elapsed,2)` seconds ( `r round(elapsed/60,2)` minutes) using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

# R environment

## R packages used

 * base R [@baseR]
 * bookdown [@bookdown]
 * data.table [@data.table]
 * ggplot2 [@ggplot2]
 * kableExtra [@kableExtra]
 * knitr [@knitr]
 * lubridate [@lubridate]
 * rmarkdown [@rmarkdown]

## Session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

# References
